\texttt{thornado}-transport: IMEX schemes for two-moment neutrino transport respecting Fermi-Dirac statistics

### 2018-10-18 Documents/M1/Astronum_2018/sections/TimeIntegration.tex
Backup the modification made in spatial discretization section.

In this section we show how to use the property we introduced at the end of the previous section to construct a constraint-preserving, implicit-explicit (IMEX) scheme.
The semi-discretization of the moment equations, Eq.~\eqref{eq:momentEquations}, results in a system of ordinary differential equations of the form
\begin{equation}
  \dfrac{d \vect{u}}{d t} = \vect{\cT}(\vect{u}) + \f{1}{\tau}\,\vect{\cQ}(\vect{u}),
\end{equation}
where
\begin{equation}
\vect{u}(t) = \left( \bcM_{1}(t),\ldots,\bcM_{N}(t)\right) ^{T}
\end{equation}
is the collection of all cell-average moments, $\vect{\cT}$ is the transport operator corresponding to the flux term, which is the first term on the right-hand side of Eq.~\eqref{eq:SemiDiscretizatedMomentEquation}, and $\f{1}{\tau}\,\vect{\cQ}$ is a linear operator corresponding to the collision term, the second term on the right-hand side of Eq.~\eqref{eq:SemiDiscretizatedMomentEquation}.

Since the realizable $\bcM$ define a convex set in $\cJ -\bcH$ space, convex-invariant schemes, which remain in their initial convex set, can be used to design realizability-preserving schemes for the two-moment model. 
The following discussion shows how to construct a convex-invariant scheme.

\subsection{Standard Implicit-Explicit Scheme}
Treating $\vect{\cT}$ explicitly and $\f{1}{\tau}\,\vect{\cQ}$ implicitly, a standard $s$-stage IMEX scheme takes the following form\cite{pareschiRusso_2005}
\begin{align}
  \vect{u}^{(i)}
  &=\vect{u}^{n}
  +\dt\sum_{j=1}^{i-1}\tilde{a}_{ij}\,\vect{\cT}(\vect{u}^{(j)})
  +\dt\sum_{j=1}^{i}a_{ij}\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(j)}),
  \quad i=1,\ldots,s, \label{imexStages} \\
  \vect{u}^{n+1}
  &=\vect{u}^{n}
  +\dt\sum_{i=1}^{s}\tilde{w}_{i}\,\vect{\cT}(\vect{u}^{(i)})
  +\dt\sum_{i=1}^{s}w_{i}\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(i)}), \label{imexIntermediate} 
\end{align}
where  $(\tilde{a}_{ij})$ and $(a_{ij})$ are coefficients of the $i$th-stage and they are the elements of matrices $\tilde{A}$ and $A$, respectively, and the vectors $\tilde{\vect{w}}=(\tilde{w}_{1},\ldots,\tilde{w}_{s})^{T}$ and $\vect{w}=(w_{1},\ldots,w_{s})^{T}$ are the weight vectors.
These coefficients and weights have to satisfy certain condition for an accurate, convergent and stable scheme.
For instance, to have second-order temporal accuracy, the following conditions are required:
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}=\sum_{i=1}^{s}w_{i}=1,
  \label{orderConditions1}
\end{equation}
and
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}\tilde{w}_{i}\,c_{i}
  =\sum_{i=1}^{s}w_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}w_{i}\,c_{i}=\f{1}{2}, 
  \label{orderConditions2}
\end{equation}
where $\tilde{c}_{i} = \sum_{j}^{s}\tilde{a}_{ij}$ and $c_{i}=\sum_{j}^{s}a_{ij}$.

>>>>

Suppose we have an algebraic closure based on Fermi-Dirac statistic (whose Eddington factor satisfies Eq.~\eqref{eq:eddingtonFactorBounds}).
We show how to construct a time integration which satisfies the bounds on moment after each time step in this section.
The semi-discretization of the moment equations, Eq.~\eqref{eq:momentEquations}, results in a system of ordinary differential equations of the form
\begin{equation}
  \dfrac{d \vect{u}}{d t} = \vect{\cT}(\vect{u}) + \f{1}{\tau}\,\vect{\cQ}(\vect{u}),
\end{equation}
where
\begin{equation}
\vect{u}(t) = \left( \bcM_{1}(t),\ldots,\bcM_{N}(t)\right) ^{T}
\end{equation}
is the collection of all cell-average moments, $\vect{\cT}$ is the transport operator corresponding to the first term on the right-hand side of Eq.~\eqref{eq:SemiDiscretizatedMomentEquation}, and $\f{1}{\tau}\,\vect{\cQ}$ corresponds to the collision operator, the second term on the right-hand side of Eq.~\eqref{eq:SemiDiscretizatedMomentEquation}.

Since the realizable $\bcM$ define a convex set in $\cJ$-$\bcH$ space, convex-invariant schemes, which remain in their initial convex set, can be used to design realizability-preserving schemes for the two-moment model. 
The following discussion shows how to construct a convex-invariant scheme.

\subsection{Standard Implicit-Explicit Scheme}
Treating $\vect{\cT}$ explicitly and $\f{1}{\tau}\,\vect{\cQ}$ implicitly, a standard $s$-stage IMEX scheme takes the following form \cite{pareschiRusso_2005}
\begin{align}
  \vect{u}^{(i)}
  &=\vect{u}^{n}
  +\dt\sum_{j=1}^{i-1}\tilde{a}_{ij}\,\vect{\cT}(\vect{u}^{(j)})
  +\dt\sum_{j=1}^{i}a_{ij}\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(j)}),
  \quad i=1,\ldots,s, \label{imexStages} \\
  \vect{u}^{n+1}
  &=\vect{u}^{n}
  +\dt\sum_{i=1}^{s}\tilde{w}_{i}\,\vect{\cT}(\vect{u}^{(i)})
  +\dt\sum_{i=1}^{s}w_{i}\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(i)}), \label{imexIntermediate} 
\end{align}
where  $(\tilde{a}_{ij})$ and $(a_{ij})$ are coefficients of the $i$-th stage and they are the elements of matrices $\tilde{A}$ and $A$, respectively.
The vectors $\tilde{\vect{w}}=(\tilde{w}_{1},\ldots,\tilde{w}_{s})^{T}$ and $\vect{w}=(w_{1},\ldots,w_{s})^{T}$ are the weight vectors.
These coefficients and weights have to satisfy certain condition for an accurate, convergent and stable scheme.
For instance, to have second-order temporal accuracy, the following conditions are required:
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}=\sum_{i=1}^{s}w_{i}=1,
  \label{orderConditions1}
\end{equation}
and
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}\tilde{w}_{i}\,c_{i}
  =\sum_{i=1}^{s}w_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}w_{i}\,c_{i}=\f{1}{2}, 
  \label{orderConditions2}
\end{equation}
where $\tilde{c}_{i} = \sum_{j}^{s}\tilde{a}_{ij}$ and $c_{i}=\sum_{j}^{s}a_{ij}$.
